---
title: "Demo of Shiny and Leaflet in R"
author: "Nick Mader"
date: "July 11, 2016"
output: html_document
runtime: shiny
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(readxl)
library(magrittr)
```

```{r data_prep, echo = FALSE}
schools <- read.csv("https://data.cityofchicago.org/api/views/2m8w-izji/rows.csv",
                    stringsAsFactors = FALSE) %>%
  within({
    temp <- Latitude
    Latitude <- Longitude # ...because these are accidentally backwards
    Longitude <- temp
    PolicyLevel <- factor(CPS.Performance.Policy.Level)
    rm(temp)
  })

# Download and merge an Excel file with enrollment data
download.file(url = "http://cps.edu/Performance/Documents/Datafiles/enrollment_20th_day_2014.xls",
              destfile = "sch_enrollment_xls.xls", method = "curl")
sch_enrollment <- read_excel("sch_enrollment_xls.xls")
sch_merge <- merge(x = schools,
                   y = sch_enrollment,
                   by.x = "School.ID",
                   by.y = "School ID") %>%
  subset(grepl("LEVEL", CPS.Performance.Policy.Level))

```

This interactive page is made interactive using R Markdown to generate the page, the Leaflet package in R, and the Shiny package in R for the user controls. Unlike the more traditional workflow of creating static reports, these tools allow us to create documents that allow our audience to interact with features, assumptions, and context that is most relevant to them, and see the results immediately. 

```{r app, echo=FALSE, fig.width = 7, fig.height = 12}

levelValues <- sort(unique(sch_merge$CPS.Performance.Policy.Level))
ui <- shinyUI(fluidPage(
  titlePanel("School Level and 5Essentials Explorer"),
  sidebarLayout(
    sidebarPanel(
      checkboxGroupInput(inputId = "myLevels",
                         label = ("School Levels to Include"),
                         choices = levelValues,
                         select  = levelValues),
      selectInput(inputId = "symbolizeField",
                  label = "Choose 5Essentials Rating to Visualie:",
                  choices = c("Effective Leaders", "Collaborative Teachers", "Supportive Environment", 
                              "Involved Family", "Ambitious Instruction"), # "Safe", "School Community", "Parent Teacher Partnership", "Quality of Facilities"
                  selected = "Involved Family"),
      sliderInput(inputId = "minResponse",
                  label = "Minimum Response Rate:",
                  min = 0.0, max = 1.0, value = 0.0, step = 0.05)
    ),
    mainPanel(
      leafletOutput("schMap")
    )
  )
))

server <- function(input, output){
  mySchs <- reactive(subset(sch_merge,
                              subset = CPS.Performance.Policy.Level %in% input$myLevels &
                                as.numeric(gsub("%", "", Student.Response.Rate ))/100 >= input$minResponse))
  essentialsLabels <- c("NOT ENOUGH DATA", "VERY WEAK", "WEAK",
                        "NEUTRAL", "STRONG", "VERY STRONG")
  symbField <- reactive(with(mySchs(),
                             factor(get(gsub(" ", "\\.", input$symbolizeField)), 
                                    levels = essentialsLabels)))
  myPalette <- colorFactor(palette = "Blues",
                           domain = essentialsLabels, 
                           ordered = TRUE)

  output$schMap <- renderLeaflet({
    leaflet(mySchs()) %>%
      addProviderTiles("CartoDB.Positron") %>%
      addCircles(lng = ~Longitude, lat = ~Latitude, popup = ~Name.of.School,
                 color = ~myPalette(symbField()), radius = ~Totals/3, stroke = FALSE,
                 opacity = 0.8, fillOpacity = 0.8) %>%
      addLegend(position = "bottomright", pal = myPalette, values = ~symbField(),
                title = input$symbolizeField)

  })
}

shinyApp(ui, server)
```


